---
title: "DataAnalysis"
author: "Marco Lee"
date: "January 23, 2020"
output: html_document
editor_options: 
  chunk_output_type: console
---
```{r, include=F}
library(dplyr)
library(ggplot2)
```

# Reading data from .csv file
```{r}
MData<-read.csv("Data/measurements.csv")
str(MData)
head(MData)
```
### Renaming columns for ease of coding
```{r}
MData<-rename(MData, spe = Species, lbW = Limb.Width, unW = unitsW, lbL = Limb.Length, unL = unitsL)
```

# Analysing data
## Standardizing units
Units are all converted to cm.
```{r}
for(i in 1:nrow(MData)){ #converting limb width
  if(MData[i,]$unW == "mm"){
    MData[i,]$lbW <- MData[i,]$lbW * 0.1
    MData[i,]$unW <- "cm"
  }
}

for(i in 1:nrow(MData)){ #converting limb length
  if(MData[i,]$unL == "mm"){
    MData[i,]$lbL <- MData[i,]$lbL * 0.1
    MData[i,]$unL <- "cm"
  }
}
str(MData)
```

## Calculating limb volume 
Limb volume is calculated by assuming that limb is a cylinder with the width as diameter and the length as height. The formula used is pi* (r^2) * h. The unit is cm^3
```{r}
MData<-mutate(MData, 
              lbV = pi*(lbW/2)^2*lbL)
head(MData)
```

```{r, eval=F, include=F}
#units not standardized in the data frame
MData<-mutate(MData, 
              lbV = ((lbW*ifelse(unW=="cm", 1, 0.1)/2)^2*pi)*
                (lbL*ifelse(unL=="cm", 1, 0.1))) #if unit is mm, first multiply by 0.1 to convert to cm
head(MData)
```

## Grouping species
Species were grouped in moths or butterflies.
```{r}
MData$spe
MData<- mutate(MData, grp=rep(NA, 10))
MData$grp<- c(1,1,1,1,1,0,0,1,1,0)
MData$grp<- ifelse(MData$grp==T, "Moth", "Butterfly")
select(MData, spe, grp)
```

## Group dimensions
Mean and standard deviation are calculated for width, length, and volume for each group.
```{r}
avg<-MData
avg<- avg %>%
  group_by(grp) %>%
  summarize(meanW=mean(lbW), SDW=sd(lbW),
           meanL=mean(lbL), SDL=sd(lbL),
           meanV=mean(lbV), SDV=sd(lbV)) %>%
  arrange(grp) #arrange in alphabetic order of group name
avg
```

#Graphs
Width is plotted against length of all species.
```{r}
ggplot(data=MData, aes(x=lbL, y=lbW)) +
  geom_point(size=I(2))+
  ylab("Limb width (cm)") +
  xlab("Limb length (cm)") +
  theme_classic() +
  theme(axis.line=element_line(colour = 'black', size = 1.2),
        axis.text=element_text(size=I(12), face="bold"),
        axis.title=element_text(size=I(15), face="bold"),
        plot.title=element_text(size=I(15), face="bold", hjust=0.5)) +
  ggtitle("Correlation between limb width and length across species")
```

### Species with above average volume
Only species with volume above the average of their respective group is included when graphing.
First, a data frame "x" with only species that fitted the criteria  is created.
```{r}
x<-MData[(MData$grp == "Moth" & MData$lbV > as.integer(avg[2,6]))|
           (MData$grp =="Butterfly" &MData$lbV > as.integer(avg[1,6])),]
x
```

Then, a scatterplot with limb width against length is created with this set of data (species with above average limb volume)
```{r}
ggplot(data=x, aes(x=lbL, y=lbW)) +
  geom_point(size=I(2))+
  ylab("Limb width (cm)") +
  xlab("Limb length (cm)") +
  theme_classic() +
  theme(axis.line=element_line(colour = 'black', size = 1.2),
        axis.text=element_text(size=I(12), face="bold"),
        axis.title=element_text(size=I(15), face="bold"),
        plot.title=element_text(size=I(15), face="bold", hjust=0.5)) +
  ggtitle("Correlation between limb width and length\nacross species with above average volume")
```

### Species with below average limb volume
The volume of species with volume below the average of their respective group is displayed in a table.

First, a data frame "y" with only species that fitted the criteria  is created.
```{r}
y<-MData[(MData$grp == "Moth" & MData$lbV < as.integer(avg[2,6]))|
           (MData$grp =="Butterfly" &MData$lbV < as.integer(avg[1,6])),]
```

Then, the columns species and limb volumn are selected and displayed.
```{r}
z<-select(MData, spe, lbV)
z<-rename(z, Species=spe, Limb.Volume=lbV) #renaming columns
z
```